To the interested reader,
\\
\\
This document is a compilation of lecture notes taken during the Spring 2023 semester for MATH3430: Ordinary Differential Equations at the University of Colorado Boulder. The course used \textit{Differential Equations and Their Applications}\footnote{Braun, M. (1993). \textit{Differential Equations and Their Applications: An Introduction to Applied Mathematics} (4th ed.). Springer. } by Martin Braun as its primary text. As such, many theorems, definitions, and content may be quoted from the book. This course was taught by Sheagan John, Ph. D.
\\
\\
The author would like to provide the following resources for students currently taking a differential equations course:
\begin{enumerate}
    \item \href{https://www.youtube.com/playlist?list=PLEC88901EBADDD980}{MIT OpenCourseWare Differential Equations Lectures From Spring 2006.}
    \item \href{https://www.youtube.com/playlist?list=PLZHQObOWTQDNPOjrT6KVlfJuKtYTftqH6}{3Blue1Brown's \textit{Overview of Differential Equations}.}
\end{enumerate}
\vphantom
\\
\\
While much effort has been put in to remove typos and mathematical errors, it is very likely that some errors, both small and large, are present. Please keep in mind that the author wrote this resource during his second semester of his undergraduate studies. If an error needs to be resolved, please contact Adithya Bhaskara at \href{mailto:adithya.bhaskara@colorado.edu}{adithya.bhaskara@colorado.edu}.
\\
\\
Finally, the author would like to dedicate this set of lecture notes to \textit{Aidan Janney}, \textit{Erika Sj\"{o}blom}, and \textit{Tate McDonald}, three of the author's closest friends who plan to take Differential Equations in the upcoming semester; Fall 2023, at the time of writing.
\\
\\
\rightline{Best Regards,}
\rightline{Adithya Bhaskara}
\vfill
\rightline{\texttt{REVISED: \today}}